---
title: "EDA"
output:
  html_document:
    df_print: paged
date: "2025-06-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(ggplot2)
library(dplyr)
library(maps)
library(viridis)
library(tidyverse)
library(DataExplorer)
library(corrplot)
library(ggpubr)
library(GGally)
library(reshape2)
library(scales)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(naniar)
#library(ggqplot)

```

```{r}
data <- read.csv("G:/STA4996_Research/data/Nutrition.csv")
```


```{r}
#head(data)
```

```{r}

colSums(is.na(data))
```


```{r}
colnames(data)[1:18] <- c("Country", "Year","UN_Region", "UN_sub_region", "Stunting_survey","Stunting_model","Underweight_survey", "Wasting_survey","Overweight_survey","Overweight_model","WHZ_sample_size","HAZ_sample_size","WAZ_sample_size","U5_population", "hcpi_a","fcpi_a","def_a","literacy_female","literacy_tot","health_expenditure","Urban_pop","rural_pop","inf_mortality","LBW","UnemploymentRate")

```
#Selected countries

```{r}

selected_countries <- c("Afghanistan", "Bangladesh", "Bolvia", "Botswana", "Brazil", "Chilie", "Colombia","Congo", "Egypt","Ethiopia", "Iran", "India", "Kenya", "Madagascar","Malaysia", "Maldives", "Mali", "Mauritius", "Mexico","Morocco", "Myanmar", "Nepal", "Nigeria", "Pakistan", "Panama", "Peru","South Africa", "Sri Lanka", "Sudan", "Thailand","Venezuela","Vietnam","Zimbabwe", "Albania", "Azerbaijan", "Tajikistan", "Turkey","Rwanda","Niger","Malawi","Mautitania","Jamaica","Indonesia","Uganda","Senegal","Burkina Faso","Burubdi","Chad","Gambia","Kuweit","Togo","Tanzania","Uruguay")
```



```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")


world$highlight <- ifelse(world$name %in% selected_countries, "Selected", "Other")


ggplot(data = world) +
  geom_sf(aes(fill = highlight)) +
  scale_fill_manual(values = c("Selected" = "tomato", "Other" = "gray90")) +
  theme_minimal() +
  labs(title = "Highlighted Countries on World Map",
       fill = "Country Type")
```



#Missing values 

```{r}
vis_miss(data)          
miss_var_summary(data)  
colMeans(is.na(data)) * 100
```
By omitting the predictor variables more than 50% of missing values, the selected socioeconomic factors are:

hcpi_a
fcpi_a
def_a
Current health expenditure(% if GDP)
urban population(% of total population)
mortality rate under 5
Low birth weight babies
unemployment rate 


# Distribution of response variables 

Stunting
```{r}
# Histograms & density
ggplot(data, aes(x = Stunting_model)) +
  geom_histogram(bins = 30, fill = "skyblue", alpha = 0.7) +
  geom_density(aes(y = ..count..), color = "darkblue") +
  theme_minimal()


```

The distribution is multi-modal (Multiple peaks)
This indicates that different countries or time periods may cluster around different stunting levels.

The highest concentrations of observations appear between 10% and 40% with several mini-peaks 

The distribution has a slightly right skew: there are some countries or years with very high stunting percentages, though most values are in the lower-to-mid range.

Long tail extending beyond 50%: high stunting in certain countries or time periods, which might act as outliers in modeling.



overweight

```{r}

ggplot(data, aes(x = Overweight_model)) +
  geom_histogram(bins = 30, fill = "salmon", alpha = 0.7) +
  geom_density(aes(y = ..count..), color = "darkred") +
  theme_minimal()
```

Right skewed

There is one clear peak around 0-2% overweight prevalence. (unimodal)
not normally distributed
the most values appear to be around 0-2%


#Correlation Analysis

```{r}
df_numeric <- data %>% 
  select(where(is.numeric)) %>% 
  drop_na()


cor_matrix <- cor(df_numeric, use = "complete.obs")
corrplot(cor_matrix, method = "color", type = "lower", tl.cex = 0.5)
```

According to the correlation plot,

Overweight modeled values are positively correlated with socioeconomic indicators literacy rate, current health expenditure, urban population and negatively correlated with the variables hcpi_a, fcpi_a, def_a, mortality rate under 5, rural population, low birth weight rate, unemployment rate 

stunting modeled values are negatively correlated with fcpi_a, hcpi_a,def_a, literacy, current health expenditure, urban population, rural population and positively correlated with mortality rate, low birth weight babies, unemployment rate 

#Outlier detction

```{r}
data %>%
  select(Country, Stunting_model, Overweight_model) %>%
  pivot_longer(cols = c(Stunting_model, Overweight_model), names_to = "indicator") %>%
  ggplot(aes(x = indicator, y = value)) +
  geom_boxplot(fill = "lightblue") +
  theme_minimal()

```
Box plots compare the distributions of overweight and stunting rates across countries.

overweight:

median is around 5-6% and the IQR is narrow
multiple outliers clustered around 15-25%,indicating several countries with unusually high overweight rates

stunting:

median is around 25-27% and the IQR is much wider 
Whiskers extend from about 2% to roughly 62%
no visible outliers, suggesting the whiskers capture the full range
more normally distributed

patterns:

most countries face low overweight but moderate-to-high stunting
few countries have severe overweight problems 
the dual burden exists but is asymmetric across the global landscape

##By country

Bar chart of average by country 

```{r}
data %>%
  group_by(Country) %>%
  summarise(across(c(Stunting_model, Overweight_model), mean, na.rm = TRUE)) %>%
  pivot_longer(cols = -Country, names_to = "indicator", values_to = "mean_val") %>%
  ggplot(aes(x = reorder(Country, mean_val), y = mean_val, fill = indicator)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  theme_minimal()

```

High prevalence in stunting than the overweight across the selected countries. 
Countries at the top (Burundi, Afghanistan... ) have high combined rates of stunting and overweight.

Dual burden pattern can be seen: some countries show significant levels of both indicators.Ex: Egypt, South Africa. This indicates the prevalence of double burden of malnutrition.

Transition pattern: As move down the chart, there is generally a shift from stunting-dominant to more balanced or overweight-dominant patterns, but this is not perfectly linear

This suggests different countries face different malnutrition challenges. 

some countries are  primarily dealing with undernutrition (stunting), others with a mixed burden. 



#Overweight

```{r}
overweight_data <- data %>%
  select(Country, Year, Overweight_model) %>%
  filter(!is.na(Overweight_model))
#overweight_data
```


```{r}

countries <- unique(overweight_data$Country)


for(country in countries) {
  country_data <- overweight_data %>% filter(Country == country)
  
  p <- ggplot(country_data, aes(x = Year, y = Overweight_model)) +
    geom_line(color = "steelblue", size = 1) +
    geom_point(color = "steelblue", size = 2) +
    labs(title = paste("Prevalence of Overweight in", country),
         x = "Year",
         y = "Overweight Prevalence (%)") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(p)
  
  
  ggsave(paste0("overweight_", gsub(" ", "_", country), ".png"), plot = p, width = 8, height = 6)
}
```

countries from the same region and the same sub-region shows different prevalence in overweight. ex: Uruguay, Bolivia, Brazil

most of the countries have declining patterns 

```{r}
data$Year <- as.numeric(data$Year)
```



```{r}
data_filtered <- data %>%
  filter(Country %in% selected_countries & !is.na(Overweight_model)) %>%
  group_by(Country) %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  select(Country, Year, Overweight_model)

```

```{r}
world_map <- map_data("world")
```


```{r}

highlighted_map <- world_map %>%
  filter(region %in% selected_countries) %>%
  left_join(data_filtered, by = c("region" = "Country"))
```

```{r}
data$
```


```{r}
library(ggplot2)
library(dplyr)
library(viridis)

# Step 1: Create a 5-year period column
highlighted_map <- highlighted_map %>%
  mutate(
    Year = as.numeric(year),  # Ensure 'year' column exists
    Period = case_when(
      Year >= 2000 & Year < 2005 ~ "2000–2004",
      Year >= 2005 & Year < 2010 ~ "2005–2009",
      Year >= 2010 & Year < 2015 ~ "2010–2014",
      Year >= 2015 & Year < 2020 ~ "2015–2019",
      Year >= 2020 ~ "2020+"
    )
  )

# Step 2: Compute average overweight by country and period
avg_overweight <- highlighted_map %>%
  group_by(country, Period) %>%
  summarise(Overweight = mean(as.numeric(Overweight_model), na.rm = TRUE)) %>%
  ungroup()

# Step 3: Join average data back to map data (to get coordinates)
map_data_period <- highlighted_map %>%
  select(long, lat, group, country, Period) %>%
  distinct() %>%
  left_join(avg_overweight, by = c("country", "Period"))

# Step 4: Plot
ggplot() +
  geom_polygon(
    data = world_map, 
    aes(x = long, y = lat, group = group),
    fill = "lightgray", 
    color = "white",
    linewidth = 0.1
  ) +
  geom_polygon(
    data = map_data_period,
    aes(x = long, y = lat, group = group, fill = Overweight),
    color = "black",
    linewidth = 0.2
  ) +
  scale_fill_viridis(
    name = "Avg Overweight %", 
    option = "plasma", 
    direction = -1, 
    na.value = "grey90",
    limits = c(0, 100)
  ) +
  coord_fixed(ratio = 1.3) +
  theme_minimal() +
  labs(
    title = "Average Overweight Percentage by Country (5-Year Periods)",
    x = "Longitude", 
    y = "Latitude"
  ) +
  facet_wrap(~ Period) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )


```



```{r}
table(highlighted_map_cleaned$Period)

```



#Stunting

```{r}
ggplot(data, aes(x = Year, y = Stunting_model, group = country, color = country)) +
  geom_line(show.legend = FALSE) +
  theme_minimal() +
  labs(title = "Stunting Trends by Country", y = "Stunting (%)")
```


```{r}

stunting_data <- data %>%
  select(Country, Year, Stunting_model) %>%
  filter(!is.na(Stunting_model))

stunting_data$Stunting <- as.numeric(stunting_data$Stunting)
stunting_data$Year <- as.numeric(as.character(stunting_data$Year))


countries <- unique(stunting_data$Country)


y_max <- max(stunting_data$Stunting, na.rm = TRUE)


for (country in countries) {
  country_data <- stunting_data %>% filter(Country == country)
  
  if (nrow(country_data) < 2) next  # Skip if not enough data

  p <- ggplot(country_data, aes(x = Year, y = Stunting_model)) +
    geom_line(color = "#E69F00", size = 1) +
    geom_point(color = "#E69F00", size = 2) +
    labs(title = paste("Prevalence of Stunting in", country),
         x = "Year",
         y = "Stunting Prevalence (%)") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_cartesian(ylim = c(0, y_max))

  print(p)

  
  safe_country <- gsub("[^A-Za-z0-9_]", "_", country)
  ggsave(filename = paste0("stunting_", safe_country, ".png"),
         plot = p, width = 8, height = 6, dpi = 300)
}
```

almost all countries have similar prevelance of stuning over time 

```{r}
data_filtered_st <- data %>%
  filter(Country %in% selected_countries & !is.na(Stunting_model)) %>%
  group_by(Country) %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  select(Country, Year, Stunting_model)
```

```{r}
world_map1 <- map_data("world")
```

```{r}
highlighted_map1 <- world_map1 %>%
  filter(region %in% selected_countries) %>%
  left_join(data_filtered_st, by = c("region" = "Country"))
```

```{r}

highlighted_map <- world_map %>%
  left_join(stunting_data, by = c("region" = "Country"))


highlighted_map$Stunting <- as.numeric(highlighted_map$Stunting)


ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray", color = "white", linewidth = 0.1
  ) +
  geom_polygon(
    data = highlighted_map,
    aes(x = long, y = lat, group = group, fill = Stunting),
    color = "black", linewidth = 0.2
  ) +
  scale_fill_viridis(
    name = "Stunting (%)",
    option = "plasma", direction = -1, na.value = "grey90",
    limits = c(0, max(highlighted_map$Stunting, na.rm = TRUE))
  ) +
  coord_fixed(ratio = 1.3) +
  theme_minimal() +
  labs(title = "Stunting Percentage by Country", x = "Longitude", y = "Latitude") +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```

Q-Q plots for stunting



```{r}
ggqqplot(data$Stunting_model, title = "QQ Plot - Stunting")

```

the stunting data is not perfectly normally distributed.


